%%%%%%%% ICML 2026 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%

\documentclass{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{multirow}
\usepackage[table,xcdraw]{xcolor}
\usepackage[dvipsnames]{xcolor}
\usepackage[normalem]{ulem}
\useunder{\uline}{\ul}{}
\usepackage{booktabs} % for professional tables
\usepackage{enumitem}

% \usepackage{algorithm}
% \usepackage{algorithmic}

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2026} with \usepackage[nohyperref]{icml2026} above.
\usepackage{hyperref}
\usepackage{pifont}


% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Use the following line for the initial blind version submitted for review:
% \usepackage{icml2026}

% For preprint, use
% \usepackage[preprint]{icml2026}

% If accepted, instead use the following line for the camera-ready submission:
\usepackage[accepted]{icml2026}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}


% if you use cleveref..
\usepackage[capitalize,noabbrev]{cleveref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% THEOREMS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\usepackage[dvipsnames]{xcolor}

% Todonotes is useful during development; simply uncomment the next line
%    and comment out the line below the next line to turn off comments
%\usepackage[disable,textsize=tiny]{todonotes}
\usepackage[textsize=tiny]{todonotes}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Z-Erase: Enabling Concept Erasure in Single-Stream Diffusion Transformers}



\definecolor{darkyellow}{RGB}{255,204,0}


\begin{document}

\twocolumn[
  \icmltitle{Z-Erase: Enabling Concept Erasure in Single-Stream Diffusion Transformers}

  % It is OKAY to include author information, even for blind submissions: the
  % style file will automatically remove it for you unless you've provided
  % the [accepted] option to the icml2026 package.

  % List of affiliations: The first argument should be a (short) identifier you
  % will use later to specify author affiliations Academic affiliations
  % should list Department, University, City, Region, Country Industry
  % affiliations should list Company, City, Region, Country

  % You can specify symbols, otherwise they are numbered in order. Ideally, you
  % should not use this facility. Affiliations will be numbered in order of
  % appearance and this is the preferred way.
  \icmlsetsymbol{equal}{*}

  \begin{icmlauthorlist}
    \icmlauthor{Nanxiang Jiang}{Beihang}
    \icmlauthor{Zhaoxin Fan}{Beihang}
    \icmlauthor{Baisen Wang}{Beihang}
    \icmlauthor{Daiheng Gao}{USTC}
    \icmlauthor{Junhang Cheng}{Beihang}
    \icmlauthor{Jifeng Guo}{Beihang}
    \icmlauthor{Yalan Qin}{Shanghai}
    \icmlauthor{Yeying Jin}{Tencent}
    \icmlauthor{Hongwei Zheng}{BAEC}
    \icmlauthor{Faguo Wu}{Beihang}
    \icmlauthor{Wenjun Wu}{Beihang}
  \end{icmlauthorlist}

  \icmlaffiliation{Beihang}{Beijing Advanced Innovation Center for Future Blockchain and Privacy Computing, School of Artificial Intelligence, Beihang University.}
  \icmlaffiliation{USTC}{University of Science and Technology of China}
  \icmlaffiliation{Shanghai}{Shanghai University}
  \icmlaffiliation{Tencent}{Tencent}
  \icmlaffiliation{BAEC}{Beijing Academy of Blockchain and Edge Computing}

  \icmlcorrespondingauthor{Zhaoxin Fan}{zhaoxinf@buaa.edu.cn}

  \icmlcorrespondingauthor{Faguo Wu}{faguo@buaa.edu.cn}

  % You may provide any keywords that you find helpful for describing your
  % paper; these are used to populate the "keywords" metadata in the PDF but
  % will not be shown in the document
  \icmlkeywords{Machine Learning, ICML}

  \vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column listing the
% affiliations and the copyright notice. The command takes one argument, which
% is text to display at the start of the footnote. The \icmlEqualContribution
% command is standard text for equal contribution. Remove it (just {}) if you
% do not need this facility.

% Use ONE of the following lines. DO NOT remove the command.
% If you have no special notice, KEEP empty braces:
\printAffiliationsAndNotice{}  % no special notice (required even if empty)
% Or, if applicable, use the standard equal contribution text:
% \printAffiliationsAndNotice{\icmlEqualContribution}

\begin{abstract}
Concept erasure is a vital safety mechanism for removing unwanted concepts from text-to-image (T2I) models. While extensively studied in U-Net and dual-stream architectures (\textit{e.g.}, Flux), this task remains under-explored in the recent emerging paradigm of single-stream diffusion transformers (\textit{e.g.}, Z-Image). In this new paradigm, text and image tokens are processed as a single unified sequence via shared parameters. Consequently, directly applying prior erasure methods typically leads to generation collapse. To bridge this gap, we introduce \textbf{Z-Erase}, the first concept erasure method tailored for single-stream T2I models. To guarantee stable image generation, Z-Erase first proposes a \textit{Stream Disentangled Concept Erasure Framework} that decouples updates and enables existing methods on single-stream models. Subsequently, within this framework, we introduce \textit{Lagrangian-Guided Adaptive Erasure Modulation}, a constrained algorithm that further balances the sensitive erasure-preservation trade-off. Moreover, we provide a rigorous convergence analysis proving that Z-Erase can converge to a Pareto stationary point. Experiments demonstrate that Z-Erase successfully overcomes the generation collapse issue, achieving state-of-the-art performance across a wide range of tasks. Code is available at: \url{https://github.com/nxjiang-jnx/Z-Erase}.
\end{abstract}

\input{sections/introduction}
\input{sections/related}
\input{sections/motivation}
\input{sections/method}
\input{sections/experiment.tex}
\input{sections/conclusion}

\section*{Acknoledgement}
This work was funded by Frontier Technologies R\&D Program of Jiangsu under Grant No. BF2025012 and by Beijing Advanced Innovation Center for Future Blockchain and Privacy Computing.

\newpage
% \clearpage
\section*{Impact Statement}
This work aims to advance the safety and reliability of modern text-to-image generative models by enabling effective concept erasure in emerging single-stream diffusion transformers. As these unified architectures are increasingly adopted due to their efficiency and performance, understanding how to responsibly control and remove undesired concepts becomes an important technical challenge. Our method provides a principled framework for mitigating the generation of harmful, sensitive, or unwanted content while preserving the general utility of the model.

We expect this work to have a positive societal impact by supporting safer deployment of generative models in real-world applications, such as content moderation, privacy protection, and compliance with ethical and legal requirements. More broadly, it contributes technical insights toward building controllable and responsible generative systems.

\newpage
\bibliography{example_paper}
\bibliographystyle{icml2026}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\appendix
\onecolumn

\input{sections/appendix}

\end{document}

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
% by Iain Murray in 2018, and modified by Alexandre Bouchard in
% 2019 and 2021 and by Csaba Szepesvari, Gang Niu and Sivan Sabato in 2022.
% Modified again in 2023 and 2024 by Sivan Sabato and Jonathan Scarlett.
% Previous contributors include Dan Roy, Lise Getoor and Tobias
% Scheffer, which was slightly modified from the 2010 version by
% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
% slightly modified from Prasad Tadepalli's 2007 version which is a
% lightly changed version of the previous year's version by Andrew
% Moore, which was in turn edited from those of Kristian Kersting and
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.
